### 2017-12-08 doc/presentation/presentation.tex
Skeleton for presentation

\documentclass[9pt]{beamer}
\usecolortheme{beaver}
\setbeamertemplate{navigation symbols}{}
\setbeamercolor{block title}{bg=white!95!black,fg=black}
\defbeamertemplate{description item}{align left}{\insertdescriptionitem\hfill}
    \setbeamertemplate{description item}[align left]
\usepackage{tikz}

\usepackage[outputdir=build]{minted}
\usepackage{tcolorbox}
\tcbuselibrary{minted,skins}

% Patch to fix https://github.com/T-F-S/tcolorbox/issues/12
\makeatletter
\def\tcb@minted@input@listing#1#2#3#4{%
  \edef\temp@a{#4}%
  \ifx\temp@a\@empty%
  \else%
    \toks@=\expandafter{#4}%
    \edef\tcb@temp{\noexpand\usemintedstyle{\the\toks@}}%
    \tcb@temp%
  \fi%
  \toks@=\expandafter{#1}%
  \edef\tcb@temp{\noexpand\inputminted[\the\toks@]}%
  \IfFileExists{\minted@outputdir#3}%
    {\tcb@temp{#2}{\minted@outputdir#3}}%
    {\tcb@temp{#2}{#3}}%
}
\makeatother

\newtcblisting{ccode}[2][]{%
  listing engine=minted,
  minted language=#2,
  minted options={breaklines,breakanywhere,fontsize=\scriptsize,gobble=8},
  listing only,
  before skip=0pt,
  after skip=8pt,
  left skip=0pt,
  right skip=0pt,
  size=fbox,
  sharp corners,
  %colframe=white!75!black,
  colframe=white,
  boxrule=0pt,
  frame hidden,
  #1
}

\newtcbinputlisting[]{\inputcode}[3][]{%
  listing engine=minted,
  minted language=#2,
  minted options={breaklines,breakanywhere,fontsize=\scriptsize,#1},
  listing file={#3},
  listing only,
  size=fbox,
  before skip=0pt,
  after skip=10pt,
  left skip=0pt,
  right skip=0pt,
  sharp corners,
  %colframe=white!75!black,
  colframe=white,
  boxrule=0pt,
  frame hidden
}

\title{Small Sparse Matrix Multiplications\\ on Knights Landing}
\author{Nathan Brei}
\institute{Technical University of Munich}
\date\today

\begin{document}
\begin{frame}
  \titlepage
\end{frame}

\begin{frame}
  \frametitle{Motivation}

  \begin{block}{SeisSol}
    Simulates earthquakes, specifically, dynamic rupture processes and seismic wave propagation over an unstructured tetrahedral mesh. 
  \end{block}

  \begin{block}{ADER-DG}
    A numerical method which uses Discontinuous Galerkin discretization in space and ADER discretization in time. It avoids a global system matrix, using local element matrix operations instead.
  \end{block}

  \begin{block}{Small Sparse Matrix Multiplications}
    The compute kernels resemble the operation shown below, applied recursively. Different sparsity patterns are used, but they are all known at compile time.
  \end{block}

  \centering
  \includegraphics[height=3cm]{images/seissol_visc.png}

\end{frame}

\begin{frame}
  \frametitle{Knights Landing}
  \begin{block}{KNL is the 2nd-generation Intel Xeon Phi many-core processor}
    \begin{itemize}
    \item A standalone processor (not just a coprocessor or GPU) which supports the entire x86 instruction set along with AVX512 extensions.
    \item Organized into a 2D grid of up to 36 tiles, each containing 2 cores and a 1MB shared L2 cache
    \item 4 threads/core $\implies$ up to 288 threads/processor
    \item 32 vector registers holding 8 doubles/register
    \item Fused-multiply-add instruction $(=:FMA)$ with mask registers and implicit broadcast
    \item 2 vpus/core, 16 double ops/vpu/cycle, 1.3e9 cycles/sec \\ $\implies$ 3 double TFlop peak performance

    \end{itemize}
  \end{block}
  \begin{block}{KNL greatly prefers large dense matrix operations to small sparse ones}
    \begin{itemize}
    \item Scalar operations limited to 1/8 peak performance
    \item Bottleneck with vector stores: 64B/cycle
    \item Bottleneck with instruction cache, decoder: min(16B/cycle, 2 instructions)
    \item \textbf{Out-of-order execution pipeline can only issue 2 instructions per cycle}
    \end{itemize}
  \end{block}
\end{frame}

\begin{frame}
  \frametitle{Past approaches used in SeisSol}
  \begin{block}{Standard compressed-sparse-column multiplication}
    \begin{itemize}
      \item[$+$] Generality.
      \item[$+$] Algorithm uses a (small) constant number of instructions.
      \item[$-$] Index lookups require indirect memory access, causing lots of problems for the execution pipeline.
      \item[$-$] $sparse \times sparse$ is fundamentally scalar. (However, BSR can be vectorized.)
      \item[$-$] Disregards the fact that the sparsity pattern is known at compile time.
    \end{itemize}
  \end{block}

  \begin{block}{Sparsity pattern unrolled into instruction stream}
    \begin{itemize}
    \item[$+$] Index lookups removed completely
    \item[$+$] $dense \times sparse$ case can be vectorized perfectly 
    \item[$-$] Generates one FMA instruction for every nonzero
    \item[$-$] (Implementation) Memory access pattern is suboptimal on KNL
    \item[$-$] (Implementation) Intel compiler struggles to generate efficient assembly
    \end{itemize}
  \end{block}
\end{frame}

\begin{frame}
  \frametitle{Present and future approaches}
  \begin{block}{Dense matrix multiplication}
    \begin{itemize}
    \item[$+$] Memory access patterns and register blockings optimized for KNL
    \item[$+$] (Implementation) Code generator emits assembly directly 
    \item[$-$] Filling in zero entries wastes both flops and memory bandwidth
    \end{itemize}
  \end{block}

  \begin{block}{My approach}
    \begin{itemize}
    \item The three existing algorithms each encounter different constraints and bottlenecks.
    \item Think of them as poles in a design space. Explore algorithms which combine their features in order to finesse around the bottlenecks.
    \item Speed up applications such as SeisSol, or demonstrate why it can't be done.
    \item Create a code generator which autotunes a sparse gemm algorithm for a reasonably arbitrary problem.
    \end{itemize}
  \end{block}
\end{frame}

\begin{frame}[fragile]
  \frametitle{The dense outer-product microkernel}
  \begin{minted}[fontsize=\footnotesize]{gas}
# Load C register block
vmovapd 0(%rdx),  %zmm16                                 # load C[:,0]
vmovapd 64(%rdx), %zmm17                                 # load C[:,1]
vmovapd 128(%rdx), %zmm18                                # load C[:,2]
# ...

# Load A register block
vmovapd 0(%rdi), %zmm0                                   # load A[:,0]
vmovapd 64(%rdi), %zmm1                                  # load A[:,1]
# ...

# Outer product of A[:,0] * B[0,:]
vfmadd231pd 0(%rsi) {1to8}, %zmm0, %zmm16                # C[:,0] += A[:,0] .* B[0,0]
vfmadd231pd 0(%rsi,%r15,1) {1to8}, %zmm0, %zmm17         # C[:,1] += A[:,0] .* B[0,1]
vfmadd231pd 0(%rsi,%r15,2) {1to8}, %zmm0, %zmm18         # C[:,2] += A[:,0] .* B[0,2]
# ...

# Outer product of A[:,1] * B[1,:]
vfmadd231pd 8(%rsi) {1to8}, %zmm1, %zmm16                # C[:,0] += A[:,1] .* B[1,0]
vfmadd231pd 8(%rsi,%r15,1) {1to8}, %zmm1, %zmm17         # C[:,0] += A[:,1] .* B[1,1]
vfmadd231pd 8(%rsi,%r15,2) {1to8}, %zmm1, %zmm18         # C[:,0] += A[:,1] .* B[1,2]
# ...
# More outer products ...

# Store C register block
vmovapd %zmm16, 0(%rdx)                                  # store C[:,0]
vmovapd %zmm17, 64(%rdx)                                 # store C[:,1]
vmovapd %zmm18, 128(%rdx)                                # store C[:,2]
# ...

  \end{minted}
\end{frame}

%\begin{frame}
%  \frametitle{Cartoon of the dense outer-product microkernel}
%\end{frame}

\begin{frame}
  \frametitle{The sparse outer-product microkernel}

  \begin{block}{Idea: Build off of the state of the art for small dense gemm}
    \begin{itemize}
    \item Assume A is dense, B is sparse
    \item Pack B into a virtual compressed-sparse-row or column format
    \item Remove all FMA and load instructions which are not needed
    \item Make this the basic building block for more complicated algorithms 
    \end{itemize}
  \end{block}
  \begin{block}{Limitations}
    \begin{itemize}
    \item Sparsity pattern must be known at compile time
    \item Pattern must fit within vector registers: $m \leq 8;  n \leq 30$
    \item (Dense $\times$ Sparse) $\neq$ (Sparse $\times$ Dense) due to vectorization
    \item Choice of row-major vs column-major format determines which of the two vectorizes nicely: $(AB)^T = B^T A^T$
    \end{itemize}
  \end{block}
\end{frame}

\begin{frame}[fragile]
  \frametitle{Fully unrolled dense x sparse}
  \begin{columns}[t]%[onlytextwidth]
    \begin{column}{0.5\textwidth}
      Generate sparse microkernels for each (k,n) block of B, and unroll these into the instruction stream.
      \begin{itemize}
      \item[$+$] Supports arbitrary matrix sizes
      \item[$+$] Dense-style register blocking for a dense $\times$ sparse algorithm.
      \item[$+$] Does not fill in any zeros -- no wasted flops

      \item[$-$] Generates an FMA instruction for each nonzero
      \item[$-$] These FMA instructions are 10B, too large to stream
      \item[$-$] Irregular location of blocks in memory prevents using tricks to reduce instruction size
      \item[$-$] Total number of nonzeros is limited by size of instruction cache.

      \end{itemize}
    \end{column}
    \begin{column}{0.5\textwidth}
      \begin{ccode}[]
        {text}
        loop over m blocks:

           unroll loop over n blocks:

              load C block into registers

              unroll loop over k blocks:

                 load A block into registers
                 microkernel for current B block
                 move A right 1 block
                 move B down 1 block

              store C block from registers
              move A to far left
              move B to top + right 1 block
              move C right 1 block

           move A down 1 block
           move C down 1 block
      \end{ccode}
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}[fragile]
  \frametitle{Performance of fully unrolled dense x sparse}
  Matrix sizes: $m=128, n=28, k=128$.
  Microkernel block sizes: $bm=8, bn=28, bk=4$.
  B is gradually filled with randomly placed nonzeros.

  \begin{columns}[t]
    \begin{column}{0.6\textwidth}
      \centering
      \includegraphics[width=6.5cm]{images/fig2.pdf}
    \end{column}
    \begin{column}{0.6\textwidth}
      \centering
      \includegraphics[width=6.5cm]{images/fig3.pdf}
    \end{column}
  \end{columns}

  \begin{itemize}
    \item For $nnz < 3100$, time-to-solution scales linearly.
    \item For $nnz > 3200$, the FMAs completely fill the instruction cache (regardless of matrix dimensions!)
    \item Fully unrolled sparse algorithm outperforms dense ``nonzero'' GFlops, but always underperforms dense ``hardware'' GFlops

  \end{itemize}
\end{frame}

\begin{frame}[fragile]
  \frametitle{Improvement: Tiled Sparse}
        Assume there exists a sparse pattern which tiles perfectly over B.
        \begin{itemize}
        \item[$+$] We can reuse a single microkernel, reducing the number of instructions
        \item[$+$] We can visit blocks in a loop efficiently, using pointer tricks to reduce instruction size.
        \item[$-$] This regularity assumption is likely too strong.
        \item[$-$] In practice, the sparsity pattern would often become dense.
        \end{itemize}

      \[
      \left[
          \begin{array}{c c c | c c c }
          0 &   &   & 10 &   &   \\
          1 & 2 &   & 11& 12&   \\
            & 3 & 4 &   & 13& 14  \\
          \hline
          5 &   &   & 15  &   &   \\
          6 & 7 &   & 16  &17   &   \\
            & 8 & 9 &   & 18  &19   \\
          \end{array}
          \right]
      \]
\end{frame}

\begin{frame}[fragile]
  \frametitle{Improvement: Block Sparse}
  Assume that B is decomposed into blocks which are either empty or `full'. `Full' blocks all have the same sparsity pattern. This is a sparse generalization of the block-sparse-column format.

  \begin{itemize}
  \item[$+$] We can reuse a single microkernel, reducing the number of instructions
  \item[$+$] The zero blocks reduce the number of filled-in nonzeros.
  \item[$-$] We can no longer visit blocks by using a loop with pointer arithmetic. \\Our options are:
    \begin{itemize}
    \item Look up the block location from a table in memory
    \item Unroll the block locations into the instruction stream and repeatedly jump to the microkernel.
    \end{itemize}
  \end{itemize}

      \[
      \left[
          \begin{array}{c c c | c c c }
          0 &   &   &    &    &    \\
          1 & 2 &   &    &    &    \\
            & 3 & 4 &    &    &    \\
          \hline
          5 &   &   & 10 &    &    \\
          6 & 7 &   & 11 & 12 &    \\
            & 8 & 9 &    & 13 & 14 \\
          \end{array}
          \right]
      \]
\end{frame}

\begin{frame}
  \frametitle{A family of dense x sparse algorithms}
  \include{flowchart}
\end{frame}

\begin{frame}[fragile]
  \frametitle{Implementation}

  \begin{block}{Code Generation}
    \begin{itemize}
    \item Python domain-specific language for directly generating assembly
    \item Includes rich assembly AST: \texttt{Operand} $\leftarrow$ \texttt{Statement} $\leftarrow$ \texttt{Block} 
    \item Higher-level forms such as loops, subroutines, and jump tables
          implemented as plain Python functions which return a \texttt{Block}
    \item Pretty-printing, analysis, simulation are implemented using Visitor Pattern
    \item Interesting theoretical problem: Moving information from runtime to compile time in a general way
    \end{itemize}
  \end{block}

  \begin{block}{Matrix Cursor}
    A high-level abstraction for traversing and examining a sparse blocked matrix using \emph{logical coordinates}, automatically translated to physical memory offsets, for a given sparsity pattern and format. 
    \begin{itemize}
    \item \verb|has_nonzero_block(pointer, block_coords) -> bool|
    \item \verb|has_nonzero_cell(pointer, cell_coords) -> bool|
    \item \verb|move(pointer, block_coords) -> (new_pointer, statement)|
    \item \verb|look(pointer, cell_coords) -> (new_pointer, memory_address)|
    \end{itemize}
    \end{block}
\end{frame}

\begin{frame}[plain,c]
\begin{center}
    \huge Thank you!
\end{center}
\end{frame}

\end{document}