### 2017-12-23 paper/appendix2.tex
Attractive code listings

\section{Appendix}
\definecolor{bg}{rgb}{0.95,0.95,0.95}

\begin{minted}[linenos,fontsize=\scriptsize,bgcolor=bg]
{julia}

const Float = Float64

"State space datatype"
mutable struct Q
    h :: Float
    hu :: Float
end
\end{minted}

\begin{minted}[linenos,fontsize=\scriptsize,firstnumber=9,bgcolor=bg]
{julia}
# Make Q a vector space

import Base.+, Base.-, Base.*

function +(lhs::Q, rhs::Q)
    Q(lhs.h + rhs.h, lhs.hu + rhs.hu)
end

function -(lhs::Q, rhs::Q)
    Q(lhs.h - rhs.h, lhs.hu - rhs.hu)
end

function *(a::Real, q::Q)
    Q(a*q.h, a*q.hu)
end
\end{minted}

\begin{minted}[linenos,fontsize=\scriptsize,firstnumber=26,bgcolor=bg]
{julia}
const G = 9.81f0

"Shallow-water point-wise flux"
function f(q :: Q)
    Q(q.hu, q.hu^2/q.h + 0.5*G*q.h^2)
end

"Shallow-water point-wise max wave speed"
function wavespeed(q :: Q)
    u = q.hu / q.h
    c = sqrt(G * q.h)
    max(abs(u-c), abs(u+c))
end

\end{minted}

\begin{minted}[linenos,fontsize=\scriptsize,firstnumber=40,bgcolor=bg]
{julia}
"Spatial domain datatype"
QS = Array{Q,1}

"Finite volume skeleton"
function run_experiment(stoptime, ncells::Int, 
    choose_dt, compute_fluxes!, apply_bcs!, apply_ics)

    qs = apply_ics(ncells)
    Fl = Array{Q,1}(ncells+2)
    Fr = Array{Q,1}(ncells+2)

    currenttime = 0
    timesteps = 0
    dx = Float(1000.0/ncells)

    while currenttime < stoptime
        apply_bcs!(qs)
        dt = choose_dt(qs, dx)
        compute_fluxes!(qs, Fl, Fr, ncells, dx, dt)
        update_qs!(qs, Fl, Fr, ncells, dx, dt)
        currenttime += dt
        timesteps += 1
    end
    return (qs, currenttime, timesteps)
end

\end{minted}
\begin{minted}[linenos,fontsize=\scriptsize,firstnumber=67,bgcolor=bg]
{julia}
# Kernels: (qs, Fl, Fr, ncells, dx, dt) -> nothing

"Lax-Friedrichs kernel"
function lxf!(qs::QS, Fl::QS, Fr::QS, 
              ncells::Int, dx::Float, dt::Float)

    a = dx/dt
    for x = 2:ncells+2
        ql,qr = qs[x-1], qs[x]
        fl,fr = f(ql), f(qr)

        Fr[x-1] = 0.5*((fr-fl) - a*(qr - ql))
        Fl[x-1] = 0.5*((fr-fl) + a*(qr - ql))
    end
end

\end{minted}
\newpage
\begin{minted}[breaklines,breakanywhere,linenos,firstnumber=85,fontsize=\scriptsize]
{julia}

"Local Lax-Friedrichs kernel"
function llxf!(qs::QS, Fl::QS, Fr::QS, 
               ncells::Int, dx::Float, dt::Float)

    for x = 2:ncells+2
        ql,qr = qs[x-1], qs[x]
        fl,fr = f(ql), f(qr)

        a = max(wavespeed(ql), wavespeed(qr))

        Fr[x-1] = 0.5*((fr-fl) - a*(qr-ql))
        Fl[x-1] = 0.5*((fr-fl) + a*(qr-ql))
    end
end

# Timestepping strategies: (QS, dx) -> dt

function make_const_dt(dt)
    return (qs, dx) -> dt
end

function cfl_dt(qs::QS, dx::Float, mu::Float)
    lambda = map(wavespeed,qs) |> maximum
    return dx / lambda * mu
end

function make_cfl_dt(correction::Float)
    return (qs, dx) -> cfl_dt(qs, dx, correction)
end

function make_lagging_dt(dt0, correction)
    dt = dt0
    function lagging_dt(qs, dx)
        result = dt
        dt = cfl_dt(qs, dx, correction)
        return result
    end
    return lagging_dt
end

# Initial conditions: ncells -> QS

function gaussian_ics(ncells::Int)
    qs = Array{Q,1}(ncells+2)
    mid = ncells / 2.0 + 1
    for x = 1:ncells+2
        pos_rel = (mid-x) / (ncells/10.0)
        qs[x] = Q(exp(-pos_rel^2) + 1, 0.0)
    end
    return qs
end

function breakingdam_ics(ncells::Int)
    [x < ncells/2+1 ? Q(2,0) : Q(1,0) for x in 1:ncells+2]
end

## Boundary conditions: QS -> nothing

function outflow_bcs!(qs::Array{Q,1})
    qs[1] = qs[2]
    qs[end] = qs[end-1]
end

function periodic_bcs!(qs::Array{Q,1})
    qs[1] = qs[end-1]
    qs[end] = qs[2]
end

function reflecting_bcs!(qs::Array{Q,1})
    qs[1].h = qs[2].h
    qs[1].hu = -qs[2].hu
    qs[end].h = qs[end-1].h
    qs[end].hu = -qs[end-1].hu
end

# Sample experiment run
qs,t,ts = run_experiment(50, 1000, make_cfl_dt(w), 
             lxf!, outflow_bcs!, breakingdam_ics)

\end{minted}